### 2016-11-28 dissertation/normal.tex
dissertation v2:part unsat update 1

\section{Formal Verification of Normal Basis Multiplier using Gr\"obner Basis}
The gate-level design of a NB multiplier can be generated using appoaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
% fig:comb_RAB

@@

One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
Analysis on experiment results shows major time consumption is on ``multi-division" part.
A matrix-based technique named as "F-4 style reduction" \cite{F4reduce} can speed up the procedure
dividing a low-degree polynomial with term-sparse polynomial ideal.

>>>>

\section{Formal Verification of Normal Basis Multiplier using Gr\"obner Basis}
The gate-level design of a NB multiplier can be generated using appoaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier}
\label{fig:comb_RAB}}
\end{figure}

@@

One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
Analysis on experiment results shows major time consumption is on ``multi-division" part.
A matrix-based technique named as "F-4 style reduction" \cite{f4} can speed up the procedure
dividing a low-degree polynomial with term-sparse polynomial ideal.

### 2016-11-28 dissertation/recha.tex
dissertation v2:part unsat update 1

Assume we are given gate-level netlist shown in figure \ref{fig:Lagrange}. It can be written as 3 
Boolean equations:
\begin{align*}
z_0 &= a_0 \lor a_1 \lor a_2 \\
z_1 &= a_1 \land \neg a_2 \lor a_0 \land \neg a_1 \land a_2 \\
z_2 &= a_1 \lor \neg a_0 \land a_2
\end{align*}
Define 2 word-level variables $A,Z$ as input and output:
$$A = \{a_2a_1a_0\},~Z = \{z_2z_1z_0\}$$
To convert bit-level to word-level, we need to find mapping $\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
or $\F_{2^3} \rightarrow \F_{2^3}$. The latter one, as mentioned in preliminaries, is a polynomial
function in $\F_{2^3}$.

For each element in $\F_{2^3}$, we write down the truth table as follows:
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
Now our objective is to abstract a function over GF $\F_{2^3}$ about word-level variables, i.e. $Z = \Func(A)$.
Recall the definition of Lagrange's interpolation:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}
The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
the real number field, the essential concept of Lagrange's interpolation remains the same. 
We can get 8 points in the affine space:
\begin{align*}
\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
\text{Point }1&: (0, 0) \gets (000,000) \\
\text{Point }2&: (1,1) \gets (001,001) \\
\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

@@

\subsection{Conventional Traversal Method}
Conceptually, the state-space of a FSM is traversed in a breadth-first
manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
The algorithm operates on the FSM $\mathcal{M} = (\sum, O, S, S^0,
\Delta, \Lambda)$ underlying a sequential circuit. In such cases, the
transition function $\Delta$ and the initial states are represented
and manipulated using Boolean representations such as BDDs or SAT
solvers. The variables $from, reached, to, new$ represent
characteristic functions of sets of states. Starting from the initial
state $from^i = S^0$, the algorithm computes the states reachable in
1-step from $from^i$ in each iteration. In line 4 of algorithm
\ref{alg:BFS}, the {\it image computation} is 
used to compute the reachable states in every execution step.

@@

\begin{Example}
For the circuit in Figure \ref{fig:fsm} (b), we have the transition
functions of the machine as:

@@

% fig:proj_reacha

@@

\subsection{Simplifying the Gr\"obner Basis Computation}
In Algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
iteration to attain the word-level polynomial representation of the next states. In practice, for 
a sequential circuit with complicated structure and large size, Gr\"obner basis computation
is intractable. To overcome the high computational complexity of computing a GB, 
we describe a method that computes a GB of a smaller subset of polynomials.
The approach draws inspirations from \cite{pruss:tcad}. 
According to proposition 2 in \cite{pruss:tcad}, if the GB is computed using {\it refined abstraction term
order} (RATO), there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
their leading monomials are not relatively prime, i.e. 
$$gcd(LM(f_w), LM(f_g)) \neq 1$$
As a well-known fact from Buchberger's algorithm, the S-polynomial 
(Spoly) pairs with relatively prime 
leading monomials will always reduce to 0 modulo the basis and have no contribution to 
the Gr\"obner basis computation.
Therefore, by removing the relatively prime polynomials from $J_{ckt}$, 
the Gr\"obner basis computation is transformed to the reduction of 
$Spoly(f_w,f_g)$ modulo $J_{ckt}$. More specifically, we turn the
GB computation into one-step multivariate polynomial division, and the obtained
remainder $r$ will only contain bit-level inputs and word-level output.

\begin{Example}
After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit 
(example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
\begin{align}
&(t_0,t_1)>(a,b,c,d)\nonumber\\&>T>(x,s_0,s_1)\nonumber
\end{align}
We can write down all polynomial generators of $J_{ckt}$:
\begin{equation*}
f_1: a+xs_0s_1+xs_0+xs_1+x+s_0s_1+s_0+s_1+1
\end{equation*}
\vspace{-0.8cm}
\begin{alignat*}{3}
& f_2: b+s_0s_1 &&~~~ f_3: c+x+xs_0 \\
& f_4: d+s_0s_1+s_0 &&~~~ f_5: t_0+ab+a+1 \\
& f_6: t_1+cd+c+d &&~~~ f_7: t_0+t_1\alpha+T 
\end{alignat*}

@@

% fig:PI

@@

Again, assume a partition $n_i~|~k$ and corresponding word-level variable is $P$. Then we can use polynomial
$P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
field}:
\begin{Theorem}
\label{thm:composite}
Let $k = m\cdot n_i$, such that $\mathbb F_{2^k} = \mathbb F_{(2^{n_i})^m}$. Let $\alpha$ be primitive root of 
$\mathbb F_{2^k}$, $\beta$ be primitive root of the ground field $\mathbb F_{2^{n_i}}$. Then
$$\beta = \alpha^\omega,~~\text{where}~~\omega = \frac{2^k-1}{2^{n_i}-1}$$
\end{Theorem}
\begin{Example}
\label{ex:PI}
In a sequential circuit, PS/NS inputs/outputs are 4-bit signals, which means we will use $\mathbb F_{2^4}$
as working field. PIs are partitioned to 2-bit vectors, which means the ground field is $\mathbb F_{2^2}$.
In ground field we can represent all possible evaluations of this PI partition $\{p_0,p_1\}$ with
$$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
Using theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
$$P = p_0 + p_1\cdot\alpha^5$$
\end{Example}
Using this method we can efficiently partition large size PIs to small number of word-level PI variables.
One limitation of this approach is PIs cannot be partitioned when $k$ is prime.

\section{Implementation of Word-level FSM Traversal Algorithm}
In this section, we will introduce the tool implementation of our approach.
The architecture of our tool consists of 4 functional components.
First, given circuit is synthesized to gate-level using state-of-art synthesizers.
Secondly, the synthesized netlist is translated to polynomial form and variables are sorted in 
RATO. This part is implemented using scripting language such as \emph{Perl}. Thirdly, 
the polynomial reduction is executed using our customized reduction engine, which is dumped in C++.
Finally, we utilize symbolic computation engine -- \textsc{Singular} \cite{DGPS} to code algorithm 
\ref{alg:refined} and execute the BFS traversal. The result is given in a univariate polynomial 
about NS word $T$, denoting the set of reachable states from given initial state.

The following example illustrates the full-blown reachability analysis approach
based on C++ and \textsc{Singular} implementation.

\begin{Example}[Full Blown Reachability Analysis Walk-through]
Usually FSM designs can be described in behavior/structural hybrid languages.
One of these languages is Berkeley logic interchange format (BLIF) \cite{BLIF},
it allows state behavior representation and logic component representation.

In this example, we use benchmark FSM ``lion9" from MCNC benchmark library.
Figure \ref{fig:BLIF} is the truth table based structural BLIF representation of this FSM.

\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/BLIF.pdf}
\caption{Sample input BLIF file}
\label{fig:BLIF}}
\end{figure}

In order to compose the polynomial set in elimination ideal, we need to synthesize it to 
gate-level netlist. Modern synthesizers including ABC \cite{brayton2010abc} and SIS \cite{SIS} 
can complete this work. Figure \ref{fig:BLIF_syn} shows a synthesized FSM given in 
BLIF format. Note that for simplifying purposes in this example, we only use 2-input logic gates.

\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/BLIF_syn.pdf}
\caption{Synthesized BLIF file}
\label{fig:BLIF_syn}}
\end{figure}

Using our interpreter, the synthesized BLIF file turns to a polynomial file customized for our 
polynomial reduction engine as figure \ref{fig:ALG} depicts. Note the first line includes all variables in RATO, the second line 
denotes the word-level variable. The third line is the irreducible polynomial of current GF $(\F_{2^4})$, the fourth line is the Spoly need to be reduced. The rest part has one polynomial in 
the elimination ideal in each line.

\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/ALG.pdf}
\caption{Polynomial file prepared for polynomial reduction engine}
\label{fig:ALG}}
\end{figure}

The result given by our reduction engine is in the format 
$$T+\Func(s_0,\dots,s_{k-1},x_0,\dots,x_{n-1})$$ 
where $s_i$ and $x_j$ denotes bit-level PS variables and PIs. Concretely in this example, the result is
\begin{align*}
T&+(\alpha^2+\alpha+1) x_0 x_1 s_1 s_3+(\alpha^3+\alpha) x_0 x_1 s_1+(\alpha+1) x_0 x_1 s_3 \\
&+(\alpha^3+1) x_0 s_1 s_3+(\alpha^3+\alpha) x_0 s_1 +(\alpha+1) x_0 s_3+(\alpha^2) x_0 \\
&+(\alpha^2+\alpha+1) x_1 s_1 s_3+(\alpha^3+\alpha) x_1 s_1+(\alpha^2+\alpha+1) x_1 s_3\\
&+(\alpha) x_1+(\alpha^3+1) s_1 s_3+(\alpha^3+\alpha) s_1+(\alpha^3+1) s_3+\alpha^2
\end{align*}
This is the transition function of this FSM. We utilized \textsc{Singular} to integrat both 
bit-word substitution and tranversal algorithm. In figure \ref{fig:SING}, ``tran" is the transition
function we just obtained. ``init\_S" is the initial state, note it equals to ``0100" register 
reset values in figure \ref{fig:BLIF}. Moreover, 2 bits PIs $x_0,x_1$ are combined to a word-level 
PI variable $P$ using our conclusion in example \ref{ex:PI}: ``def\_X" is the definition of 2-bit 
word, and ``red\_X" denotes the vanishing polynomial for word $P$.

\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/SING.pdf}
\caption{Singular script for executing bit-to-word substitution and traversal loop}
\label{fig:SING}}
\end{figure}

@@

@@

\section{Experiment Results}
\label{sec:exp_reacha}
We have implemented our traversal algorithm in 3 parts:
the first part implements polynomial reductions (division) of the
Gr\"obner basis computations, under the term order derived from the
circuit as Line 2 in Alg. \ref{alg:refined}. This is implemented 
with our customized data structure in
C++. The second part implements the bit-level to word-level
abstraction to attain transition functions at the word-level 
%$T+\mathcal{F}(S,PIs)$ 
using the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}, as Line 3 in 
Alg. \ref{alg:refined}; and the third part
executes the reachability checking iterations using 
\textsc{Singular} as well. With our
tool implementation, we have performed experiments to analyze reachability
of several FSMs. Our experiments run on a desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7-4770K Quad-core CPU, 32 GB RAM and
64-bit Ubuntu Linux OS. The experiments are shown in Table \ref{tab:recha_result}.

@@

In Table \ref{tab:recha_result}, \# States denotes the final reachable states
starting from given reset state, which given by our tool is the same with
the return value of {\it compute\_reach} in VIS. Meanwhile, from observation
of the experiment run-times, we find the reduction runtime increases
as the number of gates grows. Also, iterative reachability convergence
check's runtime reflects both the size of present state/next state
words ($k$) and the number of final reached states, which corresponds
to the degree of polynomial {\it reached} in Alg.\ref{alg:univa}. 
Although the efficiency of our initial implementation fails to compete
with the BDD based FSM analyzer VIS, the experiment
demonstrates the power of abstraction of algebraic geometry techniques
for reachability analysis applications. Currently, we are
investigating techniques that
can help us overcome the complexity of the GB computation with elimination
orders and speed-up our approach.

@@

\section{Conclusion} \label{sec:conc}
This paper has presented a new approach to perform reachability
analysis of finite state machines at the word-level. This is achieved
by modeling the transition relations and sets of states by way of
polynomials over finite fields $\Fkk$, where $k$ represents the size
of the state register bits. Subsequently using the concepts of
elimination ideals, Gr\"obner bases, and quotients of ideals, we show
that the set of reachable states can be encoded, canonically, as the
variety of a univariate polynomial. This polynomial is computed using
the Gr\"obner basis algorithm w.r.t. an elimination term
order. Experiments are conducted with a few FSMs that validate the
concept of word-level FSM traversal using algebraic geometry.

>>>>

Assume we are given gate-level netlist shown in Figure \ref{fig:Lagrange}. It can be written as 3 
Boolean equations:
\begin{align*}
z_0 &= a_0 \lor a_1 \lor a_2 \\
z_1 &= a_1 \land \neg a_2 \lor a_0 \land \neg a_1 \land a_2 \\
z_2 &= a_1 \lor \neg a_0 \land a_2
\end{align*}
Define 2 word-level variables $A,Z$ as input and output:
$$A = \{a_2a_1a_0\},~Z = \{z_2z_1z_0\}$$
To convert bit-level to word-level, we need to find mapping $\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
or $\F_{2^3} \rightarrow \F_{2^3}$. The latter one, as mentioned in preliminaries, is a polynomial
function in $\F_{2^3}$.

For each element in $\F_{2^3}$, we write down the truth table as follows:
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
Now our objective is to abstract a function over GF $\F_{2^3}$ about word-level variables, i.e. $Z = \Func(A)$.
Recall the definition of Lagrange's interpolation:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}
The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of Equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
the real number field, the essential concept of Lagrange's interpolation remains the same. 
We can get 8 points in the affine space:
\begin{align*}
\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
\text{Point }1&: (0, 0) \gets (000,000) \\
\text{Point }2&: (1,1) \gets (001,001) \\
\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

@@

\subsection{Conventional Traversal Method}
Conceptually, the state-space of a FSM is traversed in a breadth-first
manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
The algorithm operates on the FSM $\mathcal{M} = (\sum, O, S, S^0,
\Delta, \Lambda)$ underlying a sequential circuit. In such cases, the
transition function $\Delta$ and the initial states are represented
and manipulated using Boolean representations such as BDDs or SAT
solvers. The variables $from, reached, to, new$ represent
characteristic functions of sets of states. Starting from the initial
state $from^i = S^0$, the algorithm computes the states reachable in
1-step from $from^i$ in each iteration. In line 4 of Algorithm
\ref{alg:BFS}, the {\it image computation} is 
used to compute the reachable states in every execution step.

@@

\begin{Example}
For the circuit in Figure \ref{fig:fsm}(b), we have the transition
functions of the machine as:

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/proj_reacha.pdf}
\caption{Projection of the variety of circuit description ideal}
\label{fig:proj_reacha}}
\end{figure}

@@

\subsection{Simplifying the Gr\"obner Basis Computation}
In Algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
iteration to attain the word-level polynomial representation of the next states. In practice, for 
a sequential circuit with complicated structure and large size, Gr\"obner basis computation
is intractable. To overcome the high computational complexity of computing a GB, 
we describe a method that computes a GB of a smaller subset of polynomials.
The approach draws inspirations from \cite{pruss:tcad15}. 
According to proposition 2 in \cite{pruss:tcad15}, if the GB is computed using {\it refined abstraction term
order} (RATO), there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
their leading monomials are not relatively prime, i.e. 
$$gcd(LM(f_w), LM(f_g)) \neq 1$$
As a well-known fact from Buchberger's algorithm, the S-polynomial 
(Spoly) pairs with relatively prime 
leading monomials will always reduce to 0 modulo the basis and have no contribution to 
the Gr\"obner basis computation.
Therefore, by removing the relatively prime polynomials from $J_{ckt}$, 
the Gr\"obner basis computation is transformed to the reduction of 
$Spoly(f_w,f_g)$ modulo $J_{ckt}$. More specifically, we turn the
GB computation into one-step multivariate polynomial division, and the obtained
remainder $r$ will only contain bit-level inputs and word-level output.

\begin{Example}
After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit 
(Example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
\begin{align}
&(t_0,t_1)>(a,b,c,d)\nonumber\\&>T>(x,s_0,s_1)\nonumber
\end{align}
We can write down all polynomial generators of $J_{ckt}$:
\begin{equation*}
f_1: a+xs_0s_1+xs_0+xs_1+x+s_0s_1+s_0+s_1+1
\end{equation*}
\vspace{-0.8cm}
\begin{alignat*}{3}
& f_2: b+s_0s_1 &&~~~ f_3: c+x+xs_0 \\
& f_4: d+s_0s_1+s_0 &&~~~ f_5: t_0+ab+a+1 \\
& f_6: t_1+cd+c+d &&~~~ f_7: t_0+t_1\alpha+T 
\end{alignat*}

@@

\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/PI.pdf}
\caption{PI partition of a sequential circuit}
\label{fig:PI}}
\end{figure}

@@

Again, assume a partition $n_i~|~k$ and corresponding word-level variable is $P$. Then we can use polynomial
$P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
field}:
\begin{Theorem}
\label{thm:composite}
Let $k = m\cdot n_i$, such that $\mathbb F_{2^k} = \mathbb F_{(2^{n_i})^m}$. Let $\alpha$ be primitive root of 
$\mathbb F_{2^k}$, $\beta$ be primitive root of the ground field $\mathbb F_{2^{n_i}}$. Then
$$\beta = \alpha^\omega,~~\text{where}~~\omega = \frac{2^k-1}{2^{n_i}-1}$$
\end{Theorem}
\begin{Example}
\label{ex:PI}
In a sequential circuit, PS/NS inputs/outputs are 4-bit signals, which means we will use $\mathbb F_{2^4}$
as working field. PIs are partitioned to 2-bit vectors, which means the ground field is $\mathbb F_{2^2}$.
In ground field we can represent all possible evaluations of this PI partition $\{p_0,p_1\}$ with
$$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
Using Theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
$$P = p_0 + p_1\cdot\alpha^5$$
\end{Example}
Using this method we can efficiently partition large size PIs to small number of word-level PI variables.
One limitation of this approach is PIs cannot be partitioned when $k$ is prime.

\section{Implementation of Word-level FSM Traversal Algorithm}
In this section, we described the architecture of our tool which can perform word-level FSM traversal
on FSM benchmark circuits.
Briefly speaking, our tool consists of 3 functional components.
% First, the given circuit is synthesized to gate-level using state-of-art synthesizers.
First, the gate-level netlist of circuit is translated to polynomial form and variables are sorted in 
RATO. This part is implemented using scripting language such as \emph{Perl}. If 
the given benchmark is a structural/behavior hybrid description, we perform a pre-process on it to get a synthesized netlist.
Secondly, the polynomial reduction is executed using our customized reduction engine, which is written in C++.
Finally, we utilize symbolic computation engine -- \textsc{Singular} \cite{DGPS} to code Algorithm 
\ref{alg:refined} and execute the BFS traversal. The tool outputs a univariate polynomial 
for NS word $T$, denoting the set of reachable states from given initial state.

Figure \ref{fig:flowchart} illustrates the execution of reachability analysis approach
based on C++ and \textsc{Singular} implementation.

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/flowchart.pdf}
\caption{Execution process of word-level FSM traversal tool}
\label{fig:flowchart}}
\end{figure}

Usually FSM designs can be described in behavior/structural hybrid languages.
One of these languages is Berkeley logic interchange format (BLIF) \cite{BLIF},
it allows state behavior representation and logic component representation.

\begin{Example}
In this example, we use benchmark FSM ``lion9" from MCNC benchmark library.
This benchmark circuit is given as truth-table based structural BLIF representation of a FSM.

% \begin{figure}[hbt]
% \centering{
% \includegraphics[width=\textwidth]{newfig/BLIF.pdf}
% \caption{Sample input BLIF file}
% \label{fig:BLIF}}
% \end{figure}

In order to compose the polynomial set in elimination ideal, we need to synthesize it to 
gate-level netlist. Modern synthesizers including ABC \cite{brayton2010abc} and SIS \cite{SIS} 
can complete this work. In this example we use a synthesis library containing only 2-input 
AND, NAND, OR, NOR, XOR, XNOR gates as well as the inverter, the output synthesized FSM is also given in 
BLIF format.

% \begin{figure}[hbt]
% \centering{
% \includegraphics[width=\textwidth]{newfig/BLIF_syn.pdf}
% \caption{Synthesized BLIF file}
% \label{fig:BLIF_syn}}
% \end{figure}

Using our interpreter, the synthesized BLIF file turns to a polynomial file customized for our 
polynomial reduction engine. 
The input file format includes all 
variables in RATO, along with the S-poly need to be reduced and polynomials in $J_{ckt}$.
% 
% \begin{figure}[hbt]
% \centering{
% \includegraphics[width=\textwidth]{newfig/ALG.pdf}
% \caption{Polynomial file prepared for polynomial reduction engine}
% \label{fig:ALG}}
% \end{figure}

The result given by our reduction engine is in the format 
$$T+\Func(s_0,\dots,s_{k-1},x_0,\dots,x_{n-1})$$ 
where $s_i$ and $x_j$ denotes bit-level PS variables and PIs. Concretely in this example, the result is
\begin{align*}
T&+(\alpha^2+\alpha+1) x_0 x_1 s_1 s_3+(\alpha^3+\alpha) x_0 x_1 s_1+(\alpha+1) x_0 x_1 s_3 \\
&+(\alpha^3+1) x_0 s_1 s_3+(\alpha^3+\alpha) x_0 s_1 +(\alpha+1) x_0 s_3+(\alpha^2) x_0 \\
&+(\alpha^2+\alpha+1) x_1 s_1 s_3+(\alpha^3+\alpha) x_1 s_1+(\alpha^2+\alpha+1) x_1 s_3\\
&+(\alpha) x_1+(\alpha^3+1) s_1 s_3+(\alpha^3+\alpha) s_1+(\alpha^3+1) s_3+\alpha^2
\end{align*}
This is the transition function of this FSM. We utilized \textsc{Singular} to integrat both 
bit-word substitution and tranversal algorithm. In Figure \ref{fig:SING}, ``tran" is the transition
function we just obtained. ``init\_S" is the initial state, note it equals to ``0100" register 
reset values preloaded in the BLIF file. Moreover, 2 bits PIs $x_0,x_1$ are combined to a word-level 
PI variable $P$ using our conclusion in Example \ref{ex:PI}: ``def\_X" is the definition of 2-bit 
word, and ``red\_X" denotes the vanishing polynomial for word $P$.

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/SING.pdf}
\caption{Singular script for executing bit-to-word substitution and traversal loop}
\label{fig:SING}}
\end{figure}

@@

\section{Experiment Results}
\label{sec:exp_reacha}
We have implemented our traversal algorithm in 3 parts:
the first part implements polynomial reductions (division) of the
Gr\"obner basis computations, under the term order derived from the
circuit as Line 2 in Algorithm \ref{alg:refined}. This is implemented 
with our customized data structure in
C++. The second part implements the bit-level to word-level
abstraction to attain transition functions at the word-level 
%$T+\mathcal{F}(S,PIs)$ 
using the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}, as Line 3 in 
Algorithm \ref{alg:refined}; and the third part
executes the reachability checking iterations using 
\textsc{Singular} as well. With our
tool implementation, we have performed experiments to analyze reachability
of several FSMs. Our experiments run on a desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7-4770K Quad-core CPU, 32 GB RAM and
64-bit Ubuntu Linux OS. The experiments are shown in Table \ref{tab:recha_result}.

@@

In Table \ref{tab:recha_result}, \# States denotes the final reachable states
starting from given reset state, which given by our tool is the same with
the return value of {\it compute\_reach} in VIS. Meanwhile, from observation
of the experiment run-times, we find the reduction runtime increases
as the number of gates grows. Also, iterative reachability convergence
check's runtime reflects both the size of present state/next state
words ($k$) and the number of final reached states, which corresponds
to the degree of polynomial {\it reached} in Algorithm \ref{alg:univa}. 
Although the efficiency of our initial implementation fails to compete
with the BDD based FSM analyzer VIS, the experiment
demonstrates the power of abstraction of algebraic geometry techniques
for reachability analysis applications. Currently, we are
investigating techniques that
can help us overcome the complexity of the GB computation with elimination
orders and speed-up our approach.

@@

\section{Conclusion} \label{sec:conc}
This dissertation has presented a new approach to perform reachability
analysis of finite state machines at the word-level. This is achieved
by modeling the transition relations and sets of states by way of
polynomials over finite fields $\Fkk$, where $k$ represents the size
of the state register bits. Subsequently using the concepts of
elimination ideals, Gr\"obner bases, and quotients of ideals, we show
that the set of reachable states can be encoded, canonically, as the
variety of a univariate polynomial. This polynomial is computed using
the Gr\"obner basis algorithm w.r.t. an elimination term
order. Experiments are conducted with a few FSMs that validate the
concept of word-level FSM traversal using algebraic geometry.
